---
title: Analysis of parasitoid performance vs transfected lineages of hamiltonella-protected
  aphids
date: ""
editor_options:
  chunk_output_type: inline
output:
  pdf_document:
    toc: yes
  html_document:
    theme: yeti
    toc: yes
    toc_float: yes
authors: Maxime & Youn
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
```


# Packages

```{r packages-loading}
#[R v4.0.5]

if(require("pacman")==T){
  library("pacman")
}else{
  install.packages("pacman")
}

if(require("cmdstanr")==T){
  library("cmdstanr")
}else{
  install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
}


pacman::p_load(cmdstanr, brms, Rmisc, tidyverse, patchwork, ggplot2, ggtext, bayesplot, tidybayes, kableExtra)

#library(rstan)
## Stan backend

#library(cmdstanr)  # [github::stan-dev/cmdstanr] v0.3.0 
## Stan backend (another)

#library(brms)      # CRAN v2.15.0        
## the interface we are using

#library(tidyverse) # CRAN v1.3.0   

#plotting
#library(patchwork) # CRAN v1.1.1   
#library(ggtext)    # CRAN v0.1.1      

#library(bayesplot) # CRAN v1.8.0   
#library(tidybayes) # CRAN v2.3.1   

library(here)      # CRAN v1.0.1
## tool for workspace location management
## automatically detects the workspace after loading the package if Rstudio was opened using the script file (or with new R project)
## load last to avoid conflicts

## some useful default settings
# rstan_options(auto_write = TRUE) #for rstan
options(mc.cores = 4) ## reduce/increase depending on cores available
```

# Overview

Loading the data

```{r data-load}

raw_data=read.table(here("data","data_parasitism_R.txt"),h=T, dec=".")
head(raw_data)
raw_data=subset(raw_data, raw_data$Aphid_line!="TR-0191")

```

Take a look at the dataset

```{r overlook}
## tinkering better order of wasps for plotting
order.wasps=c("Yp20-027",	"Yp20-038", "Yp20-081", "Yp20-004", "Yp20-006",	"Yp20-009", "Yp20-013", "Yp20-035", "Yp20-037", "Yp20-050", "Yp20-058", "Yp20-060", "Yp20-068", "Yp20-075",	"Yp20-079",	 "Yp20-084", "F316", "G339", "G363", "G392", "S251", "S350", "Yp20-003", "Yp20-032",	"Yp20-002", "Yp20-021", "Yp20-033", "Yp20-044", "Yp20-048", "Yp20-055", "Yp20-063", "Yp20-070", "Yp20-071", "Yp20-072", "Yp20-080")
order.aphids=c("NTR-407","TR-0016","TR-0103","TR-0131","TR-0619","TR-0042","TR-0076","TR-0214","TR-0244")
order.origin=c("Aphis_urticata","Aphis_ruborum","Aphis_fabae_fabae","Aphis_hederae")

data<-na.omit(raw_data) %>% 
  mutate(Parasitoid_line=factor(Parasitoid_line),
         Aphid_line=factor(Aphid_line),
         Hamiltonella_host_species=factor(Hamiltonella_host_species),
         Wasp_host_species=factor(Wasp_host_species)) %>% 
  mutate(replicate_fullID=paste(Aphid_line,Replicate),
         observation=1:length(Aphid_line))

data=data %>%
  mutate(Parasitoid_line = factor(Parasitoid_line, levels = order.wasps)) %>%
  arrange(Parasitoid_line)    
data=data %>%
  mutate(Aphid_line = factor(Aphid_line, levels = order.aphids)) %>%
  arrange(Aphid_line)
data=data %>%
  mutate(Hamiltonella_host_species = factor(Hamiltonella_host_species, levels = c("none",order.origin))) %>%
  arrange(Hamiltonella_host_species) 

success=as.numeric(data$Mummies_amount)/as.numeric(data$Nymphs_corrected)
data$Parasitism_success=success
mean_data=group.CI(Parasitism_success ~ Parasitoid_line*Aphid_line, data=data, ci = 0.95)
mean_data$Wasp_host_species=rep(c(
  "Aphis_urticata","Aphis_urticata","Aphis_urticata","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae"
),9)


## bubble plot
colors=c("#66ccff", "#aa80ff", "#88ff4d", "#ff9900")
ggSize=0.25
bubbles<-ggplot(data=NULL, aes(Parasitoid_line, Aphid_line))+
  geom_point(data=mean_data, aes(size=Parasitism_success.mean*25, color=Wasp_host_species, shape=16))+ ##adjust size to frequency
  scale_shape_identity()+
  scale_colour_manual(values=colors)+ ##define custom colors for protected and unprotected
  geom_text (data=mean_data[mean_data$Parasitism_success.mean>=0.01,], aes(label=round(Parasitism_success.mean*100,0)), size=3.5)+ 
  scale_size_identity()+ ##do not plot zero frequency
  scale_x_discrete(limits= levels(mean_data$Parasitoid_line))+ ##reorder axes according to levels of factor
  scale_y_discrete(labels=levels(mean_data$Aphid_line))+
  labs(title=NULL, x="Parasitoid line", y="Aphid line")+ theme(legend.position="none")+
  ##geom_hline(yintercept=1.5,lwd=ggSize)+
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position="none",
        axis.title.y=element_text(size=10),
        axis.title.x=element_text(size=10, vjust=-1),
        axis.text.x=element_text(angle=90,size=10,colour='black'),
        axis.text.y=element_text(size=10,colour='black'),
        axis.ticks=element_line(size=ggSize), 
        axis.line = element_line(size =ggSize, colour='black'))
bubbles

if(file.exists(here("R_output","Plot_bubbles.pdf"))){
  
  }else
    {
pdf(height=4, width=10, file=here("R_output","Plot_bubbles.pdf"))
bubbles
dev.off()}
```

```{r data exploration plot #1, parasitism success relative to control}
data_control=subset(data, Hamiltonella_host_species=="none")
mean10=group.CI(Parasitism_success ~ Aphid_line, data=data_control, ci = 0.95)
mean10=mean10 %>%
  mutate(Aphid_line = factor(Aphid_line, levels = order.aphids)) %>%
  arrange(Aphid_line)

mean11=group.CI(Parasitism_success ~ Aphid_line, data=data, ci = 0.95)
mean11=mean11 %>%
  mutate(Aphid_line = factor(Aphid_line, levels = order.aphids)) %>%
  arrange(Aphid_line)

mean11$Parasitism_success.mean.norm=mean11$Parasitism_success.mean-mean10$Parasitism_success.mean
mean11$Parasitism_success.upper.norm=mean11$Parasitism_success.upper-mean10$Parasitism_success.mean
mean11$Parasitism_success.lower.norm=mean11$Parasitism_success.lower-mean10$Parasitism_success.mean
data$Parasitism_success_centered=data$Parasitism_success-mean10$Parasitism_success.mean 

if(file.exists(here("R_output","Parasitism_relative_to_control.pdf"))){
 colors=c("black", "#ff9900","#ff9900","#ff9900","#88ff4d","#66ccff","#66ccff","#aa80ff","#aa80ff")
par(mar=c(10,5,1,1)+.1)
stripchart(data$Parasitism_success_centered~ data$Aphid_line, vertical = TRUE, 
           method = "jitter", jitter=0.2, add = F, pch = 20, cex=0.5, col=adjustcolor(colors, alpha.f=0.4), xlab="", ylab="Parasitism success\n (% difference to control)", ylim=c(-0.5,0.8), las=2)
points(mean11$Parasitism_success.mean.norm~mean11$Aphid_line,col="black", pch=20, cex=1)
segments(x0=c(1:10),x1=c(1:10),y0=mean11$Parasitism_success.mean.norm,y1=mean11$Parasitism_success.upper.norm,col="black")
segments(x0=c(1:10),x1=c(1:10),y0=mean11$Parasitism_success.mean.norm,y1=mean11$Parasitism_success.lower.norm,col="black")
abline(mean11$Parasitism_success.upper.norm[1],0, col="red", lty=2)
abline(mean11$Parasitism_success.lower.norm[1],0, col="red", lty=2)
 }else
    {
pdf(height=4, width=4, file=here("R_output","Parasitism_relative_to_control.pdf"))
colors=c("black", "#ff9900","#ff9900","#ff9900","#88ff4d","#66ccff","#66ccff","#aa80ff","#aa80ff")
par(mar=c(5,5,1,1)+.1)
stripchart(data$Parasitism_success_centered~ data$Aphid_line, vertical = TRUE, 
           method = "jitter", jitter=0.2, add = F, pch = 20, cex=0.5, col=adjustcolor(colors, alpha.f=0.4), xlab="", ylab="Parasitism success\n (% difference to control)", ylim=c(-0.5,0.8), las=2)
points(mean11$Parasitism_success.mean.norm~mean11$Aphid_line,col="black", pch=20, cex=1)
segments(x0=c(1:10),x1=c(1:10),y0=mean11$Parasitism_success.mean.norm,y1=mean11$Parasitism_success.upper.norm,col="black")
segments(x0=c(1:10),x1=c(1:10),y0=mean11$Parasitism_success.mean.norm,y1=mean11$Parasitism_success.lower.norm,col="black")
abline(mean11$Parasitism_success.upper.norm[1],0, col="red", lty=2)
abline(mean11$Parasitism_success.lower.norm[1],0, col="red", lty=2)

dev.off()}


```
```{r data exploration plot #2 Paraisitism success local vs non-local}
data_no_control=subset(data, Hamiltonella_host_species!="none")
data_no_control$Hamiltonella_host_species=droplevels(data_no_control)$Hamiltonella_host_species
data_no_control=data_no_control %>%
  mutate(Wasp_host_species = factor(Wasp_host_species, levels = order.origin)) %>%
  arrange(Wasp_host_species)
data_matching_origin=subset(data_no_control, as.factor(data_no_control$Wasp_host_species)==as.factor(data_no_control$Hamiltonella_host_species)) %>%
  mutate(Wasp_host_species = factor(Wasp_host_species, levels = order.origin)) %>%
  arrange(Wasp_host_species)

data_unmatching_origin=subset(data_no_control, as.factor(data_no_control$Wasp_host_species)!=as.factor(data_no_control$Hamiltonella_host_species)) %>%
  mutate(Wasp_host_species = factor(Wasp_host_species, levels = order.origin)) %>%
  arrange(Wasp_host_species)

mean6=group.CI(Parasitism_success ~ Parasitoid_line, data=data_control, ci = 0.95)
mean6$Wasp_host_species=rep(c(
  "Aphis_urticata","Aphis_urticata","Aphis_urticata","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_ruborum","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_fabae_fabae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae","Aphis_hederae"
),1)
mean8=group.CI(Parasitism_success ~ Parasitoid_line, data=data_matching_origin, ci = 0.95)
mean9=group.CI(Parasitism_success ~ Parasitoid_line, data=data_unmatching_origin, ci = 0.95)

if(file.exists(here("R_output","Parasitism_local_vs_nonlocal.pdf"))){
plot(NULL, xlim=c(0:1), ylim=c(0:1), xlab="Parasitism success\n vs non-local H. defensa", ylab="Parasitism success\n vs local H. defensa"
)
palette(c("#00a0f0","#5c0aff","#53c71a","#e68a00"))
segments(x0=mean9$Parasitism_success.mean,x1=mean9$Parasitism_success.mean,y0=mean8$Parasitism_success.lower,y1=mean8$Parasitism_success.upper,col=as.factor(mean6$Wasp_host_species))
segments(x0=mean9$Parasitism_success.upper,x1=mean9$Parasitism_success.lower,y0=mean8$Parasitism_success.mean,y1=mean8$Parasitism_success.mean,col=as.factor(mean6$Wasp_host_species))
points(mean8$Parasitism_success.mean~mean9$Parasitism_success.mean,
       pch = 20, cex=1, col=as.factor(mean6$Wasp_host_species))
abline(0,1, col="red", lty=3)
}else
    {
pdf(height=4, width=4, file=here("R_output","Parasitism_local_vs_nonlocal.pdf"))
par(mar=c(5,5,1,1)+.1)
plot(NULL, xlim=c(0:1), ylim=c(0:1), xlab="Parasitism success\n vs non-local H. defensa", ylab="Parasitism success\n vs local H. defensa"
)
palette(c("#00a0f0","#5c0aff","#53c71a","#e68a00"))
segments(x0=mean9$Parasitism_success.mean,x1=mean9$Parasitism_success.mean,y0=mean8$Parasitism_success.lower,y1=mean8$Parasitism_success.upper,col=as.factor(mean6$Wasp_host_species))
segments(x0=mean9$Parasitism_success.upper,x1=mean9$Parasitism_success.lower,y0=mean8$Parasitism_success.mean,y1=mean8$Parasitism_success.mean,col=as.factor(mean6$Wasp_host_species))
points(mean8$Parasitism_success.mean~mean9$Parasitism_success.mean,
       pch = 20, cex=1, col=as.factor(mean6$Wasp_host_species))
abline(0,1, col="red", lty=3)

dev.off()}

```

# Data preparation

```{r data preparation}
data<-na.omit(raw_data) %>% 
  mutate(Parasitoid_line=factor(Parasitoid_line),
         Aphid_line=factor(Aphid_line),
         Hamiltonella_host_species=factor(Hamiltonella_host_species),
         Wasp_host_species=factor(Wasp_host_species)) %>% 
  mutate(replicate_fullID=paste(Aphid_line,Replicate),
         observation=1:length(Aphid_line))

data<-data %>% 
  ## create origin columns with no underscore for goood looking plots
  ## and species names we will be able to display 
  ## in italic using the markdown facilities in ggtext
  mutate(
    Origin_para=str_replace_all(Wasp_host_species,"_"," "),
    Origin_ham=str_replace_all(Hamiltonella_host_species,"_"," ")
  ) %>% 
  mutate(
    Origin_para=str_replace_all(Origin_para,"Aphis","A."),
    Origin_ham=str_replace_all(Origin_ham,"Aphis","A.")
  ) %>% 
  mutate(Origin_para=factor(paste("Parasitoid source host: *",Origin_para,"*",sep="")),
         Origin_ham=paste("*",Origin_ham,"*",sep="")) %>%
  mutate(Origin_ham=fct_recode(Origin_ham,none="*none*")) %>% 
  ## we finish by ensuring the no-Ham control is the first level for both 
  ## Aphid_line and host_origin variables, for the model
  mutate(Aphid_line = fct_relevel(Aphid_line, 
                             "NTR-407", after=0),  
         Hamiltonella_host_species=fct_relevel(factor(Hamiltonella_host_species),
                                               "none",after=0),
         Origin_ham=fct_relevel(factor(Origin_ham),
                                 "none",after=0)
         )


```


# Models with bayesian approach

## Priors

Weakly informative priors from McElreath (2020)
```{r prior}
prior=c(
  set_prior("normal(0,1)", class="b"), # can be improved
  set_prior("normal(0,1.5)",class="b", # specific priors for parameters interpretable as the logit of an actual proportion
            # i.e. wasp-Origin-specific intercepts
            coef=c("Wasp_host_speciesAphis_fabae_fabae",
                   "Wasp_host_speciesAphis_hederae",
                   "Wasp_host_speciesAphis_ruborum",
                   "Wasp_host_speciesAphis_urticata")),
  set_prior("normal(0,1)", class="sd"), # can be replaced by exponential(1) (larger)
  set_prior("lkj(2)",class="cor") ##correlations between random effects
  # can loosen that with lkj(1)
)

prior_nocors=c(
  set_prior("normal(0,1.5)", class="b"),
  set_prior("normal(0,1)", class="sd")
)
```

## Model 1

Let's start with a simple binomial model:

```{r mod_noOLRE}
if(file.exists(here("R_output","mod_noOLRE.Rdata"))){
  # this if-else statement is avoid re-fitting a model if there is already one existing in R_output
  # to override, re-run the model and re-save manually by selecting only the relevant code lines (or delete the Rdata object before relaunching this code chunk)
  load(here("R_output","mod_noOLRE.Rdata"))
  }else
    {

mod <- brm(
  bf(Mummies_amount|trials(Nymphs_corrected)~
       0+Wasp_host_species+Wasp_host_species:Hamiltonella_host_species+ ## origin in the wild are fixed effects
       (Hamiltonella_host_species|Parasitoid_line)+ ## parasitoid lines have a random effect nested by hamiltonella origin
       (0+Aphid_line|Parasitoid_line)+ ## hamiltonella lines have a random effect nested by parasitoid origin
       (1|replicate_fullID), ## not changing much
       #(1|Block) + ## block effect is not valid since it's confounded with parasitoid line
     family="binomial"),
  data=data,
  chains=4, iter=2000, seed=666,
  prior=prior,
  backend="cmdstanr" ## better with "cmdstanr", but "rstan" does the job as well
)

save(list=c("mod"), file=here("R_output","mod_noOLRE.Rdata"))
}

```

Let's check the fitting of the model and prediction vs observed distribution:

```{r pp_check}
pp_check(mod)
pp_check(mod,"stat_2d")
```

Not good fit of predictions, clear overdispersion.
We also have too low prediction of 0, but this is more likely the result of the overdispersion than actual zero-inflation


## Model 2: with OLRE

Let's try with OLRE:

```{r mod}
if(file.exists(here("R_output","mod_OLRE.Rdata"))){
  load(here("R_output","mod_OLRE.Rdata"))
  }else
    {
modOLRE <- brm(
  bf(Mummies_amount|trials(Nymphs_corrected)~
       0+Wasp_host_species+Wasp_host_species:Hamiltonella_host_species+ ## origin in the wild are fixed effects
       (Aphid_line|Wasp_genotype)+ ## hamiltonella lines have a random effect nested by parasitoid genotype
       
       ####CHECK IF CORRECT ORDER OF NESTED EFFECTS
       
       
       (1|replicate_fullID)+ ## random effect of the replicate inside each not changing much
       #(1|Block) #+ ## block effect is not valid since it's confounded with parasitoid line
       (1|observation), ## this is the OLRE
     family="binomial"),
  data=data,
  chains=4, iter=2000, seed=666,
  prior=prior,
  ### control=list(adapt_delta=0.9),
  ### https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup
  ### https://mc-stan.org/users/documentation/case-studies/divergences_and_bias.html
  ###
  backend="cmdstanr" ## better with "cmdstanr", but rtan does the job as well
)


save(list=c("modOLRE"), file=here("R_output","mod_OLRE.Rdata"))
}

```

```{r pp_check_bis}
pp_check(modOLRE)
pp_check(modOLRE,"stat_2d")
```
Ok, that's better

# Plots and postprocessing

## Plots

Let's plot data and make comparisons based on the OLRE model

Predictions base on fixed effects only:

```{r fits_fix}
fits<-data %>% 
  select(Hamiltonella_host_species,Wasp_host_species,
         Origin_ham,Origin_para) %>% 
  distinct() %>% 
  mutate(Nymphs_corrected=1) %>% 
  add_fitted_draws(modOLRE,re_formula = NA) %>%
  ungroup()

```

Table of observed parasitism success means of parasitoid lines x aphids lines combinations

```{r lines_means}
obsmeans<-data %>% 
  group_by(Hamiltonella_host_species,
           Wasp_host_species,
           Origin_ham,Origin_para,
           Wasp_genotype,Aphid_line) %>% 
  summarise(mean_obs=mean(Mummies_amount/Nymphs_corrected,na.rm=TRUE),
            sd_obs=sd(Mummies_amount/Nymphs_corrected,na.rm=TRUE),
            N_obs=sum(Nymphs_corrected)) %>% 
  ungroup()
obsmeans2<-subset(obsmeans, obsmeans$Aphid_line!="TR-0619")
```

Plot

```{r plot_v1}
data=data %>%
  mutate(Aphid_line = factor(Aphid_line, levels = order.aphids)) %>%
  arrange(Aphid_line)
data=data %>%
  mutate(Origin_ham = factor(Origin_ham, levels = c("none", "*A. urticata*", "*A. ruborum*", "*A. fabae fabae*", "*A. hederae*"))) %>%
  arrange(Origin_ham) 

plot_v1<-ggplot(fits)+
  geom_boxplot(data=data,
              aes(Origin_ham,
                  Mummies_amount/Nymphs_corrected,
                  group=paste(Aphid_line,Wasp_genotype),
                  color=Aphid_line),
              alpha=0.3)+
  stat_eye(aes(Origin_ham,.value),
           normalize="xy",
           .width=c(0.001,0.95),slab_alpha=0.5)+
  scale_x_discrete("*Hamiltonella* source")+
  scale_y_continuous("Proportion of nymphs parasitized")+
  facet_wrap(~Origin_para) +
  theme(
        #legend.position = "none",
        strip.text = element_markdown(),
        axis.text.x = element_markdown(),
        axis.title.x= element_markdown())
plot_v1+scale_color_manual(values = c("black","#ffc266", "#e68a00", "#995c00", "#53c71a", "#4dc3ff", "#00a0f0", "#9966ff", "#5c0aff"))
```

```{r plot_v2}
obsmeans=obsmeans %>%
  mutate(Origin_ham = factor(Origin_ham, levels = c("none", "*A. urticata*", "*A. ruborum*", "*A. fabae fabae*", "*A. hederae*"))) %>%
  arrange(Origin_ham)

plot_v2<-ggplot(fits)+
  geom_jitter(data=obsmeans,
              aes(Origin_ham,
                  mean_obs, color=Origin_ham),
              height=0,width=0.25,
              size=2,alpha=0.7)+
  stat_eye(aes(Origin_ham,.value),
           normalize="xy",
           .width=c(0.001,0.95),slab_alpha=0.5)+
  scale_x_discrete("*Hamiltonella* source")+
  scale_y_continuous("Proportion of nymphs parasitized")+
  facet_wrap(~Origin_para) +
  theme(
        #legend.position = "none",
        strip.text = element_markdown(),
        axis.text.x = element_markdown(),
        axis.title.x= element_markdown())

plot_v2+scale_color_manual(values = c("black","#e68a00","#53c71a","#00a0f0","#5c0aff"))
```


Re-do fits, but scaled to the parasitism success of each wasp line against the defense-free aphids

```{r fits_fix}
translate=data %>% 
  select(Origin_ham,Origin_para,Hamiltonella_host_species,Wasp_host_species) %>% 
  distinct()

fits_scaled<-data %>% 
  select(Hamiltonella_host_species,Wasp_host_species,
         Origin_ham,Origin_para) %>% 
  distinct() %>% 
  mutate(Nymphs_corrected=1) %>% 
  add_fitted_draws(modOLRE,re_formula = NA, scale="linear") %>%
  ungroup() %>% 
  group_by(Wasp_host_species) %>% 
  compare_levels(.value,by=Hamiltonella_host_species, comparison = "control") %>%
  mutate(Hamiltonella_host_species= str_remove(Hamiltonella_host_species, " - none"))%>% 
  left_join(translate)

fits_scaled_probs<-fits_scaled %>% 
  mutate(.value= (exp(.value)/(1+exp(.value))))

```

Scaled plot

```{r plot_scaled}
plot_scaled<-ggplot(fits_scaled)+
  stat_eye(aes(Origin_ham,.value,fill = Origin_ham),
           normalize="xy",
           .width=c(0.001,0.95),slab_alpha=0.5)+
  geom_hline(yintercept = 0, lty=2)+
  scale_x_discrete("*Hamiltonella* source")+
  scale_y_continuous("Relative parasitism success (logit scale)")+
  facet_wrap(~Origin_para) +
  theme(
        #legend.position = "none",
        strip.text = element_markdown(),
        axis.text.x = element_markdown(),
        axis.title.x= element_markdown())
plot_scaled+
scale_fill_manual(values = c("#e68a00","#53c71a","#00a0f0","#5c0aff"))+
theme(legend.position = "none")
```
```{r plot_scaled_probs}
plot_scaled_probs<-ggplot(fits_scaled_probs)+
  stat_eye(aes(Origin_ham,.value,fill=Origin_ham),
           normalize="xy",
           .width=c(0.001,0.95),slab_alpha=0.5)+
  geom_hline(yintercept = 0.5, lty=2)+
  scale_x_discrete("*Hamiltonella* source")+
  scale_y_continuous("Relative parasitism success")+
  facet_wrap(~Origin_para) +
  theme(
        #legend.position = "none",
        strip.text = element_markdown(),
        axis.text.x = element_markdown(),
        axis.title.x= element_markdown())
plot_scaled_probs+
scale_fill_manual(values = c("#e68a00","#53c71a","#00a0f0","#5c0aff"))+
theme(legend.position = "none")
```

## Pairwise comparisons


Extract the modeled pairwise differences between treatments/variables

```{r pairwise_diff}
## Differences between parasitoid origins (host), by hamiltonella origin (original host)
fits %>% 
  mutate(.iteration=.draw) %>% 
  group_by(Hamiltonella_host_species) %>% 
  compare_levels(.value,by=Wasp_host_species) %>% 
  mean_hdi(.value) %>%   ## here with mean, but it also works with mean_hdi, median_qi, mean_qi
  mutate(.value=round(.value, digits = 2)) %>% 
  mutate(.value = cell_spec(.value, color=ifelse(.value > 0 & .lower>0, "green", ifelse(.value<0 & .upper<0, "red", "black"))))%>% ## color postive and negative values that do not overlap the zero
    kbl(booktabs = T, linesep = "", escape=FALSE) %>%
  kable_paper("striped",full_width = F) %>%
  scroll_box(width = "100%", height = "300px")



## Differences between hamiltonella origins, par parasitoid origin
fits %>% 
  mutate(.iteration=.draw) %>% 
  group_by(Wasp_host_species) %>% 
  compare_levels(.value,by=Hamiltonella_host_species) %>% 
  mean_hdi(.value) %>%   ## here with mean, but it also works with mean_hdi, median_qi, mean_qi
  mutate(.value=round(.value, digits = 2)) %>% 
  mutate(.value = cell_spec(.value, color=ifelse(.value > 0 & .lower>0, "green", ifelse(.value<0 & .upper<0, "red", "black"))))%>% 
    kbl(booktabs = T, linesep = "", escape=FALSE) %>%
  kable_paper("striped",full_width = F) %>%
  scroll_box(width = "100%", height = "300px")



## All pairwise differences between all combinations of hamiltonella origin x parasitoid origin
fits %>% 
  mutate(.iteration=.draw) %>% 
  mutate(combo=interaction(Wasp_host_species,Hamiltonella_host_species)) %>% 
  compare_levels(.value,by=combo) %>% 
  mean_hdi(.value) %>%   ## here with mean, but it also works with mean_hdi, median_qi, mean_qi
  mutate(.value=round(.value, digits = 2)) %>% 
  mutate(.value = cell_spec(.value, color=ifelse(.value > 0 & .lower>0, "green", ifelse(.value<0 & .upper<0, "red", "black"))))%>% 
    kbl(booktabs = T, linesep = "", escape=FALSE) %>%
  kable_paper("striped",full_width = F) %>%
  scroll_box(width = "100%", height = "300px")


```

## Pairwise comparisons, but grouping by other variables than fixed effects (random effects)

```{r fits_RE}
fitsRE<-data %>% 
  select(Hamiltonella_host_species,Wasp_host_species,
         Origin_ham,Origin_para,
         Aphid_line,Wasp_genotype) %>% ## we include "lineage" variables for aphids (hamiltonella) and parasitoids
  distinct() %>% 
  mutate(Nymphs_corrected=1) %>% 
  add_fitted_draws(modOLRE,
                   re_formula = ~(Aphid_line|Wasp_genotype)) %>% # we do not include random effects at a level below lineages
  ungroup()
```

```{r}
left_join(obsmeans,fitsRE) %>% 
  ggplot()+
  stat_pointinterval(aes(mean_obs,.value),
                     .width=c(0.001,0.95))+
  scale_x_continuous("observed mean proportion")+
  scale_y_continuous("predicted mean proportion")+
  geom_abline(intercept=0,slope=1)
```


```{r}
fitsRE %>% 
  mutate(.iteration=.draw) %>% 
  group_by(Aphid_line) %>% 
  compare_levels(.value,by=Wasp_genotype) %>% 
  mean_hdi(.value) %>%   ## here with mean, but it also works with mean_hdi, median_qi, mean_qi
  mutate(.value=round(.value, digits = 2)) %>% 
  mutate(.value = cell_spec(.value, color=ifelse(.value > 0 & .lower>0, "green", ifelse(.value<0 & .upper<0, "red", "black"))))%>% 
    kbl(booktabs = T, linesep = "", escape=FALSE) %>%
  kable_paper("striped",full_width = F) %>%
  scroll_box(width = "100%", height = "300px")
```

# variance partitioning: a quick look

OK so this is a quick look on how much genotype pair-level variance is due to genotype origin vs. anything else about the genotype. It's on the *logit scale* (for how to do on the observed % scale, see: doi:10.1534/genetics.115.186536)(for how to partition the totality of the variance in the model, including within-genotype variance, see e.g. doi:10.1111/j.2041-210x.2012.00261.x and follow up publications by the same)

we first need to estimate the variance explained by fixed effects. Simple: we make predictions based on fixed effects only, then estimate their variance. Important: we make these predictions using every point of the original data (no "use distinct() to keep only one unique value" stuff)

```{r fits_fix_logit}
fits_logit<-data %>% 
  select(Hamiltonella_host_species,Wasp_host_species,
         Origin_ham,Origin_para) %>% 
  mutate(Nymphs_corrected=1) %>% 
  add_fitted_draws(modOLRE,re_formula = NA,scale="linear") %>%
  ungroup()


var_fix<- fits_logit %>% 
  group_by(.draw) %>% 
  summarise(VF=var(.value))
```

Now we do the same with the variance explained by line-level random effects. If there were only random intercepts, it would be as simple as extracting the relevant variance parameter (the SDs you see in the summary). But we have random slopes (the wasp genotype-specific effects of each Hamiltonella). So the easier way (we can check later if is strictly right), is to simply do the same as above, but including these random effects in the predictions:

```{r fits_linelevel_logit}
fits_logit2<-data %>% 
  select(Hamiltonella_host_species,Wasp_host_species,
         Origin_ham,Origin_para, 
         Aphid_line, Wasp_genotype) %>% 
  mutate(Nymphs_corrected=1) %>% 
  add_fitted_draws(modOLRE,re_formula = ~(Aphid_line|Wasp_genotype),scale="linear") %>%
  ungroup()


var_partition<- fits_logit2 %>% 
  group_by(.draw) %>% 
  summarise(Vline=var(.value)) %>% 
  left_join(var_fix)
```

that's it! we can plot it

```{r plot_varpart}

ggplot(var_partition)+
  geom_halfeyeh(aes(VF/Vline),.width=c(0.001,0.95))+
  scale_x_continuous("proportion of line-level variation explained by line origin hosts (logit scale)")

```

